---
title: "p8105_hw2_ss6692"
output: github_document
date: "2023-09-28"
---

## Problem 2

```{r libraries, message=FALSE}
library(tidyverse)
library(readxl)
library(dplyr)
library(janitor)
library(tidyr)
```

### Loading and cleaning the Mr. Trash Wheel dataset

```{r clean_mr_trashwheel}
mr_trashweel = 
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", range = "A2:N550", sheet = "Mr. Trash Wheel")|>
  janitor::clean_names() |>
  mutate(trashwheel_name = "mr_trashwheel")|>
  mutate(homes_powered = (weight_tons * 500)/30)|>
  select(-month, -year) |>
  separate(date, into = c("year", "month", "day"), convert = TRUE) |>
  drop_na()
```

### Loading and cleaning the Professor Trash Wheel dataset

```{r clean_prof_trashwheel}
prof_trashwheel = 
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", range = "A2:M97", sheet = "Professor Trash Wheel")|>
  janitor::clean_names() |>
  mutate(trashwheel_name = "professor") |>
  mutate(homes_powered = (weight_tons * 500)/30)|>
  select(-month, -year) |>
  separate(date, into = c("year", "month", "day"), convert = TRUE) |>
  drop_na()
```

### Loading and cleaning the Gwynnda Trash Wheel dataset

```{r clean_gwynnda}
gwynnda = 
  readxl::read_excel("data/202207 Trash Wheel Collection Data.xlsx", range = "A2:K110", sheet = "Gwynnda Trash Wheel")|>
  janitor::clean_names() |>
  mutate(trashwheel_name = "gwynnda")|>
  mutate(homes_powered = (weight_tons * 500)/30)|>
  select(-month, -year) |>
  separate(date, into = c("year", "month", "day"), convert = TRUE) |>
  drop_na()
```

### Merging the three datasets

```{r trashwheel_merged}
trashwheel_tidy = 
  bind_rows(prof_trashwheel, gwynnda, mr_trashweel) |>
  select(trashwheel_name, everything())
```

```{r gwynnda_cigs, echo=FALSE}
  gwynnda_cigs =  
  filter(gwynnda, month == "7", year == 2021) |> 
  summarise(total_cigarette_butts = sum(cigarette_butts))
```

Using `bind_rows`, I merged the datasets `mr_trashwheel`, `prof_trashwheel` and `gwynnda` into a cleaned dataset titled `trashwheel_tidy`. In `trashwheel_tidy`, there are `r nrow(trashwheel_tidy)` observations and `r ncol(trashwheel_tidy)` columns. The dataset explores how much trash was collected by the trash wheels using the variable `weight_tons,` the types of trash collected, and the number of `homes_powered` using the collected trash.Using this dataset, we know that Gwynnda Trash Wheel collected `r gwynnda_cigs$total_cigarette_butts` cigarette butts. Additionally, we know that Professor Trash Wheel collected `r sum(prof_trashwheel$weight_tons)` tons of trash.
 


## Problem 3
### Importing and cleaning baseline data
```{r baseline_data}
baseline_data = read_csv(file = "data/MCI_baseline.csv") |>
  janitor::row_to_names(row_number = 1) |>
  janitor::clean_names() |>
  mutate(sex = ifelse(sex == "0", "male", "female")) |>
  mutate(apoe4 = ifelse(apoe4 == "0", "non_carrier", "carrier")) |>
  filter(age_at_onset != ".")
  baseline_data$current_age = as.numeric(baseline_data$current_age)
```

```{r prop_female, echo=FALSE}
prop_female =
  filter(baseline_data, sex == "female", apoe4 == "carrier") |> 
  nrow() / sum(baseline_data$sex == "female")*100
```

I imported the baseline data using `read_csv`. I used `janitor::row_to_names` to prompt the dataset to ignore the variable descriptions and use row 2 as the column names. Then, I changed `sex` and `apoe4` so they are no longer numeric and more descriptive of the variable. Finally, I used the `filter`function to remove participants who had no MCI at baseline. Of the `r nrow(read_csv("data/MCI_baseline.csv"))-1` participants, `r nrow(baseline_data)` develped MCI and remained in the cleaned dataset. The average baseline age is `r mean(baseline_data$current_age)` years. In this dataset, `r prop_female`% of the females are APOE4 carriers.

### Importing and cleaning amyloid data
```{r amyloid_data}
amyloid_data = read_csv(file = "data/mci_amyloid.csv") |>
  janitor::row_to_names(row_number = 1) |>
  janitor::clean_names() |>
  rename(id = study_id)
```

The amyloid data was imported and cleaned similarly to the baseline data. There are `r nrow(amyloid_data)` observations and `r ncol(amyloid_data)` variables. Therefore, there are 4 more participants in the amyloid dataset than in the baseline dataset. I renamed `study_id` as `id` so we can merge the two datasets in the next step.  

### Merging datasets by participants

```{r mci_data}
mci_data = 
  inner_join(baseline_data, amyloid_data, by = "id")
```

I merged the data by `id` number so only the participants who appear in both datasets are present. In this dataset, there are `r nrow(mci_data)` participants.

### Exporting `mci_data` as a .csv

```{r}
write.csv(mci_data, file = "data/mci_data.csv")
```

